### 2024-10-11 manuscript.tex
Update .github/workflows/ai-revision.yaml [ci skip]
This build is based on
https://github.com/pivlab/annual_review_of_biomedical_data_science/commit/d940251e024223111145d42fb373b9cc191fe474.

This commit was created by the following CI build and job:
https://github.com/pivlab/annual_review_of_biomedical_data_science/commit/d940251e024223111145d42fb373b9cc191fe474/checks
https://github.com/pivlab/annual_review_of_biomedical_data_science/actions/runs/11285114313

This manuscript
(\href{https://pivlab.github.io/annual_review_of_biomedical_data_science/v/6f7af112710ae5f8179f2e9d88b13533a8cab962/}{permalink})
was automatically generated
from \href{https://github.com/pivlab/annual_review_of_biomedical_data_science/tree/6f7af112710ae5f8179f2e9d88b13533a8cab962}{pivlab/annual\_review\_of\_biomedical\_data\_science@6f7af11}
on October 11, 2024.

@@

Approaches based on gene modules have the potential to go beyond prior knowledge and capture patterns that might be unique to different disease datasets.
First, although a gene module can be aligned to a pathway, this does not mean that it is restricted to prior knowledge as in standard pathway analyses.
Being aligned with pathways means that a module resembles a known mechanism, and this helps in interpretability and also in separating from other patterns that might be related to technical noise but the module could capture other genes as well that are not part of the pathway but potentially also involved in the function.
Second, gene modules can also be extracted from very large, heterogeneous datasets such as recount2, or more specific ones such as the Human Trisome Project, which includes gene expression from people with Down Syndrome (DS).
Approaches such as PLIER and PhenoPLIER have been fundamental to extract DS-specific gene modules related to obesity, a common co-occurring condition The omni-PLIER framework, combining the PLIER model with the omnigenic approach, uncovers molecular mechanisms linking Down syndrome (DS) and obesity, identifying key latent variables with top genes such as GPX1 and MCL1, which mediate the relationship between trisomy 21 and body mass index (BMI), offering insights into DS-specific and general obesity-related processes (\protect\hyperlink{ref-B7FF0NeZ}{95}).

>>>>

This manuscript
(\href{https://pivlab.github.io/annual_review_of_biomedical_data_science/v/d940251e024223111145d42fb373b9cc191fe474/}{permalink})
was automatically generated
from \href{https://github.com/pivlab/annual_review_of_biomedical_data_science/tree/d940251e024223111145d42fb373b9cc191fe474}{pivlab/annual\_review\_of\_biomedical\_data\_science@d940251}
on October 11, 2024.

@@

Approaches based on gene modules have the potential to go beyond prior knowledge and capture patterns that might be unique to different disease datasets.
First, although a gene module can be aligned to a pathway, this does not mean that it is restricted to prior knowledge as in standard pathway analyses.
Being aligned with pathways means that a module resembles a known mechanism, and this helps in interpretability and also in separating from other patterns that might be related to technical noise but the module could capture other genes as well that are not part of the pathway but potentially also involved in the function.
Second, gene modules can also be extracted from very large, heterogeneous datasets such as recount2, or more specific ones such as the Human Trisome Project, which includes gene expression from people with Down Syndrome (DS).
Approaches such as PLIER and PhenoPLIER have been fundamental to extract DS-specific gene modules related to obesity, a common co-occurring condition (\protect\hyperlink{ref-B7FF0NeZ}{95}).
